---
title: "p8105_hw2_kw2873"
author: "Keyi Wang"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)

```
### problem 1
```{r}
### basic manipulation for dataset
Wheel_data = 
   readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = 1,
                      range = "A2:N408" ) %>%
   janitor::clean_names() %>%
   drop_na(dumpster) %>%
   mutate(
   sports_balls = as.integer(sports_balls) 
   )
```


```{r}
### clean and modify data set for precipitation in 2017
preci_2017_data =
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = 4,range = "A2:B14") %>%
  drop_na() %>%
  janitor::clean_names()  %>%
  mutate( year = 2017) 


### clean and modify dataset for precipitation in 2018
preci_2018_data =
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                     sheet = 4,range = "A2:B14") %>%
  drop_na() %>%
  janitor::clean_names()  %>%
  mutate( year = 2018) 

### combiing dataset for precipitation in 2017 and 2018 
preci_17_18_data =
   bind_rows(preci_2017_data, preci_2018_data) %>%
   mutate(month = month.name[month]
   )
```
There are `r nrow(Wheel_data)`observations in Wheel_data dataset, and `r nrow(preci_17_18_data)` observations in preci_17_18_data. Wheel_data dataset contains data of`r nrow(Wheel_data)` dumpsters from`r pull(Wheel_data, date)[1]` to `r pull(Wheel_data, date)[nrow(Wheel_data)]`. The total precipitation in 2018 is `r sum(pull(preci_2018_data, total))`, and the total precipitation in 2017 is `r sum(pull(preci_2017_data, total))`. The median number of sports balls in a dumpster in 2017 is 
`r median(pull(Wheel_data %>% filter(year == 2017), sports_balls))`

### problem 2
```{r }
### manipulation pols monts data
pols_months_data = 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year","month","day")) %>%
  pivot_longer(
  starts_with("prez"),
  names_to = "president",
   names_prefix = "prez_",
   values_to = "value") %>%
   filter(!(value == "0")) %>%   
   select(year, month, everything(), -value, -day) %>%
   mutate(year = as.character(year), month = month.name[as.numeric(month)])


### manipulation of snp.csv
snp_data = 
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, into = c("day","month","year"),sep = "/") %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  select(year,month, day, close) %>%
  arrange(year,month)


### manipulation of unemployment.csv
unemployment_data = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
     cols = jan:dec,
     names_to = "month",
     values_to = "umemployment value"
     ) %>%
 mutate(year = as.character(year), month = replace(month, values = month.name),
        )

### joining dataset
merged_data = 
  left_join(pols_months_data, snp_data, by = c("year","month")) %>% 
  left_join(., unemployment_data, by = c("year","month"))
```
In pols_months_data, there are `r nrow(pols_months_data)` observations of `r ncol(pols_months_data)` variables related to the number of national politicians who are democratic or republic at any given time. There are date varaibles, number of republican or democratic governors/senators/representatives/president on the associated date. The range of dates is from 1949 to 2015. Some of the key variables are president and year. 

In the dataset of snp_data, there are `r nrow(snp_data)` observations of `r ncol(snp_data)` variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. There are year, month and close values of the S&P stock index on the associated date. The range of dates is from 1950 to 2015. Some of the key variables are close, year and month. 
In the dataset called unemployment_data, there are `r nrow(unemployment_data)` observations of `r ncol(unemployment_data)` variables related to the precentage of unemployment on the associated dates, which is pretty self-explanatory. The range of dates is from 

In the dataset of merged_data, this is pretty much the same with what was stated above, because this dataset is just a combination of all three datasets above. There are `r nrow(merged_data)` observations of `r ncol(merged_data)` variables.

### problem 3
```{r }

baby_names_data = 
  read_csv("./data/Popular_Baby_Names.csv", col_types = "icccii") %>% 
  janitor::clean_names() %>%
  mutate_all(~replace(., ethnicity == "ASIAN AND PACI", 
                      "ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate_all(~replace(., ethnicity == "BLACK NON HISP", 
                      "BLACK NON HISPANIC")) %>% 
  mutate_all(~replace(., ethnicity == "WHITE NON HISP", 
                      "WHITE NON HISPANIC")) %>% 
 distinct() ### remove replicate row

Olivia_female =
  filter(baby_names_data, childs_first_name == "Olivia", gender == "FEMALE") %>% 
  pivot_wider(
    id_cols = c("ethnicity", "year_of_birth"),
    names_from = "year_of_birth",
    values_from = "rank")
  


Popular_name_male = 
  filter(baby_names_data, rank == 1, gender == "MALE") %>% 
  pivot_wider(
    # find unique identifer
    id_cols = c("ethnicity", "year_of_birth"),
    # put years into column
    names_from = "year_of_birth",
    # put names into table cells
    values_from = "childs_first_name"
  )

white_non_hisp_2016 = 
  filter(baby_names_data, year_of_birth == 2016, gender == "MALE", ethnicity == "WHITE NON HISPANIC") %>% 
  mutate(count = as.integer(count), rank = as.integer(rank))

white_non_hisp_2016

ggplot(data = white_non_hisp_2016) + geom_point(aes(x = rank, y = count))


```


